### 2012-05-02 inst/examples/policycosts/stochastic_norms.Rmd
single source for norms

`ro fig.width=10, fig.height=7, message=FALSE, comment=NA, cache=FALSE, verbose=TRUE or`

``` {r setup, include=FALSE, cache=FALSE}
opts_knit$set(upload.fun = socialR::flickr.url)
options(device = function(width=5, height=5) pdf(NULL, width = width, height = height))
````

# Asymmetric Policy Costs 
 * author Carl Boettiger, <cboettig@gmail.com>
 * license: CC0

## Setup the system

``` {r setupsource}
rm(list=ls())   
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)

delta <- 0.05     # economic discounting rate
OptTime <- 50     # stopping time
gridsize <- 50   # gridsize (discretized population)
sigma_g <- 0.2    # Noise in population growth
sigma_m <- 0.     # noise in stock assessment measurement
sigma_i <- 0.     # noise in implementation of the quota
reward <- 0       # bonus for satisfying the boundary condition

## @knitr noise_dists
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() rlnorm(1,  0, sigma_m) # mean 1
z_i <- function() rlnorm(1,  0, sigma_i) # mean 1

f <- BevHolt                # Select the state equation
pars <- c(1.5, 0.05)             # parameters for the state equation
K <- (pars[1] - 1)/pars[2]  # Carrying capacity (for reference 
xT <- 0                     # boundary conditions
x0 <- K

## @knitr profit_
profit <- profit_harvest(price = 10, c0 = 30, c1 = 10)

## @knitr create_grid_
x_grid <- seq(0.01, 1.2 * K, length = gridsize)  
h_grid <- seq(0.01, 0.8 * K, length = gridsize)

## @knitr reed_sdp
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g )
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime, xT, 
                     profit, delta, reward=reward)
````

``` {r fees}
L1 <- function(c2) function(h, h_prev)  c2 * abs(h - h_prev) 
free_increase <- function(c2) function(h, h_prev)  c2 * abs(min(h - h_prev, 0)) # increasing harvest is free
free_decrease <- function(c2) function(h, h_prev)  c2 * max(h - h_prev, 0) # decreasing harvest is free
fixed <-  function(c2) function(h, h_prev) c2
L2 <- function(c2) function(h, h_prev)  c2 * (h - h_prev) ^ 2
````

Solve the policy cost for the specified penalty function

``` {r policycost_optim}
c2 <- 4
penalty <- free_decrease(c2)
policycost <- optim_policy(SDP_Mat, x_grid, h_grid, OptTime, xT, 
                    profit, delta, reward, penalty = penalty)
cache=FALSE
````

### Simulate

Now we'll simulate 100 replicates of this stochastic process under the optimal harvest policy determined above.  We use a modified simulation function that can simulate an alternate policy (the Reed optimum, where policy costs are zero, `opt$D` ) and a focal policy, `policycost$D`

``` {r simulate_policy, cache=cache}
sims <- lapply(1:100, function(i)
  simulate_optim(f, pars, x_grid, h_grid, x0, policycost$D, z_g, z_m, z_i, opt$D, profit=profit, penalty=penalty)
  )
````

Make data tidy (melt), fast (data.tables), and nicely labeled.

``` {r tidy, cache=cache}
dat <- melt(sims, id=names(sims[[1]]))  
dt <- data.table(dat)
setnames(dt, "L1", "reps") # names are nice
````

### Plots

A single replicate, alternate dynamics show the Reed optimum, while harvest/fishstock show the impact of having policy costs.

``` {r rep1, cache=cache}
ggplot(subset(dt,reps==1)) +
  geom_line(aes(time, alternate)) +
  geom_line(aes(time, fishstock), col="darkblue") +
  geom_line(aes(time, harvest), col="purple") + 
  geom_line(aes(time, harvest_alt), col="darkgreen") 
````

A second replicate

``` {r rep2, cache=cache}
ggplot(subset(dt,reps==2)) +
  geom_line(aes(time, alternate)) +
  geom_line(aes(time, fishstock), col="darkblue") +
  geom_line(aes(time, harvest), col="purple") + 
  geom_line(aes(time, harvest_alt), col="darkgreen") 
````

Profits

``` {r rep1profit, cache=cache}
ggplot(subset(dt,reps==1)) +
  geom_line(aes(time, profit_fishing)) +
  geom_line(aes(time, policy_cost), col="darkblue") 
````

We can visualize the equilibrium policy for each possible harvest:

``` {r policy, cache=cache}
policy <- sapply(1:length(h_grid), function(i) policycost$D[[i]][,1])
ggplot(melt(policy)) + 
  geom_point(aes(h_grid[Var2], (x_grid[Var1]), col=h_grid[value]-h_grid[Var2])) + 
    labs(x = "prev harvest", y = "fishstock") +
      scale_colour_gradientn(colours = rainbow(4)) 
````

Here we plot previous harvest against the recommended harvest, coloring by stocksize.  Note this swaps the y axis from above with the color density.  Hence each x-axis value has all possible colors, but they map down onto a subset of optimal harvest values (depending on their stock).

``` {r harvestchanges, cache=cache}
policy <- sapply(1:length(h_grid), function(i) policycost$D[[i]][,1])
ggplot(melt(policy)) + 
  geom_point(aes(h_grid[Var2], (h_grid[value]), col = x_grid[Var1]), position=position_jitter(w=.005,h=.005), alpha=.5) + 
    labs(x = "prev harvest", y = "harvest") +
      scale_colour_gradientn(colours = rainbow(4)) 
````