### 2012-08-23 inst/examples/unmanaged_warning.Rmd
step by step: initial signals for #21

`ro tidy=FALSE, warning=FALSE, comment=NA, message=FALSE, verbose=TRUE, cache=TRUE, cache.path="unmanaged_warning/" or`

``` {r  echo=FALSE, cache=FALSE}
options(device = function(width = 5, height = 5) {
    pdf(NULL, width = width, height = height)
})
````

# Detecting warning signals in an unmanaged system

This is a precursor exercise to managed_warning.Rmd to calibrate a reasonable set of parameters over which to evaluate the detection.

``` {r  setup, echo=FALSE, cache=FALSE}
# load required libraries
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
rm(list=ls())
````

Simulate from the May population dynamics function

``` {r  simulate }
n <- 5000
x <- vector(mode="double", length=n)
a <- x
x[1] <- 8 # positive equilibrium
z <- rlnorm(n, 0, .1)
r = .75; k = 10; a[1]=1.55; H=1; Q = 3
for(t in 1:n){
  x[t+1] = z[t] *  x[t] * exp(r * (1 - x[t] / k) - a[t] * x[t] ^ (Q - 1) / (x[t] ^ Q + H ^ Q)) 
  a[t+1] = a[t] + .001
}
````

### Plot of timeseries

``` {r  p0}
plot(x, type='l')
````

Truncate the timeseries

``` {r truncate}
y <- x[x > 1.5]
````

### Calculate warning signals on the truncated series.

``` {r libs2}
library(earlywarning)
dat <- data.frame(time=1:length(y), value=y)
````

``` {r earlywarning}
acor_tau <- warningtrend(dat, window_autocorr)
var_tau <- warningtrend(dat, window_var)
````

## Model based statistics

Fit the models

``` {r modelbased}
A <- stability_model(dat, "OU")
B <- stability_model(dat, "LSN")
observed <- -2 * (logLik(A) - logLik(B))
m <- B$pars["m"]
````