### 2012-11-22 inst/examples/mult_uncertainty_test.Rmd
numerical instabilities in uniform noise case?

Test multiple uncertainty function

```{r  setup, echo=FALSE, cache=FALSE}
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
````

```{r plotting, include=FALSE}
theme_publish <- theme_set(theme_bw(14))
theme_publish <- theme_update(
        # panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        legend.title=element_blank(), 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        plot.background = element_rect(fill = "transparent",colour = NA))
```

## Current version of the multiple uncertainty function

```{r}
#' SDP under multiple uncertainty
#'
#' Computes the SDP solution under the case of growth noise, 
#' implementation errors in harvesting, and meaurement errors in 
#' the stock assessment.  
#'
#' @param f the growth function of the escapement population (x-h)
#'   should be a function of f(t, y, p), with parameters p
#' @param p the parameters of the growth function
#' @param x_grid the discrete values allowed for the population size, x
#' @param h_grid the discrete values of harvest levels to optimize over
#' @param sigma is the shape parameters for noise distribution (sigma_g, sigma_m, sigma_i) (default is no noise)
#' @param pdfn is the probability density function (same functional form is used for growth, measure, implement).  (Default is uniform)
#' @param profit is the profit function (defaults to the realized harvest)
#' @return The D matrix giving the optimal harvest for each possible state in each timestep.  
#' Hence the optimal policy at time t is given by the policy function D[,t].  
#' Values of D[,t] correspond to the index of h_grid.  Indices of of D[,t] correspond to states in y_grid.  
#' @export
SDP_multiple_uncertainty <- function(f, p, x_grid, h_grid, Tmax = 25,
                                     sigmas =c(sigma_g=0.3, sigma_m=0, sigma_i=0), 
                                     pdfn = pdfn, profit = function(x,h) pmin(x, h)){

    sigma_g <- sigmas['sigma_g']      
    sigma_m <- sigmas['sigma_m']
    sigma_i <- sigmas['sigma_i']
    n_x <- length(x_grid)
    n_h <- length(h_grid)
    D <- matrix(NA, nrow=n_x, ncol=Tmax)

    ## Compute the transition matrices 
    P <-  outer(x_grid, h_grid, profit) 
    M <- rownorm( outer(x_grid, x_grid, pdfn, sigma_m) )
    I <- rownorm( outer(h_grid, h_grid, pdfn, sigma_i) )

    # integrate over uncertainty in I and X 
    F <- lapply(1:n_h, function(q){  
      t(sapply(1:n_x, function(y){
        out <- numeric(n_x)
        mu <- sum(sapply(1:n_x, function(x)
          f(x_grid[x],h_grid,p) %*% I[q,] *  M[y,x])) # Implementation & Measurement error
        if(mu==0)
          out[1] <- 1   
        else {
         out <- pdfn(x_grid, mu, sigma_g)
         if(sum(out) == 0)
           out[1] = 1
         out
        }
        out/sum(out)
      }))
    })

    Ep <- M %*% P %*% t(I)
    V <- Ep
    for(t in 1:Tmax){
      D[,(Tmax-t+1)] <- apply(V, 1, which.max)
      v_t <- apply(V, 1, max) # vector of current values
      V <- sapply(1:n_h, function(j){ # updated value matrix
        Ep[,j] + (1-delta) * M %*% F[[j]] %*% v_t
      })
    }
    list(D=D, M=M, I=I, P=P, Ep=Ep, V=V, F=F)
}

# row-normalize the probability distribution
rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
# binning routine (vectorized)
snap_to_grid <- function(x, grid) sapply(x, function(x) grid[which.min(abs(grid - x))])   
```

``` {r stateeq}
f <- function(x, h, p){
  A <- p[1] 
  B <- p[2] 
  s <- pmax(x-h, 0)
  A * s/(1 + B * s)
}
profit = function(x,h) pmin(x, h)

pars <- c(1.5, 0.05)
K <- (pars[1] - 1)/pars[2]
xmin <- 0
xmax <- 1.5 * K
n_x <- 200
n_h <- n_x
x_grid <- seq(xmin, xmax, length = n_x)  
h_grid <- seq(xmin, xmax, length = n_h)
delta <- 0.05
xT <- 0
OptTime <- 25
sigma_g <- 0.5
````

Old school way:

``` {r  SDP_Mat}
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g)
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime, xT, profit, delta, reward=0)
````

New way:

```{r}
# Uniform pdf where width scales with mean (probably better way to do that)
# And handles degenerate/delta fn cases such as no width and zero mean
FUN <- function(P, mu, s){
  if(mu == 0){
    as.integer(P == 0)
  } else if(s > 0){
    if(mu > 0) dunif(P, mu * (1 - s), mu * (1 + s))
  }
  else { # delta spike
    P <- snap_to_grid(P, x_grid)
    mu <- snap_to_grid(mu, x_grid)
    as.integer(P == mu)
  }
}
pdfn <- Vectorize(FUN)
```

```{r mult_uncertainty}
out <- SDP_multiple_uncertainty(f, pars, x_grid, h_grid, OptTime, sigmas = c(sigma_g=0.3, sigma_m=0.3, sigma_i=0.3), pdfn=pdfn)  
```

```{r noisetypes}
new <- SDP_multiple_uncertainty(f, pars, x_grid, h_grid, OptTime, sigmas = c(sigma_g=0.3, sigma_m=0, sigma_i=0), pdfn=pdfn)  
m <- SDP_multiple_uncertainty(f, pars, x_grid, h_grid, OptTime, sigmas = c(sigma_g=0.3, sigma_m=0.3, sigma_i=0), pdfn=pdfn)  
i <- SDP_multiple_uncertainty(f, pars, x_grid, h_grid, OptTime, sigmas = c(sigma_g=0.3, sigma_m=0, sigma_i=0.3), pdfn=pdfn)  
```

Plot the policy function (in terms of escapement, `x-h`, rather than harvest `h`) at equilibrium (first time-step):

```{r policyfn_data}
require(reshape2)
policies <- melt(data.frame(stock=x_grid, old = x_grid[opt$D[,1]], new = x_grid[new$D[,1]], meas = x_grid[m$D[,1]], imp = x_grid[i$D[,1]]), id="stock")
```

```{r policyfunctions}
q1 <- ggplot(policies, aes(stock, stock - value, color=variable)) + geom_point() + xlab("stock size") + ylab("escapement") 
q1
```