### 2012-11-27 inst/examples/policycosts/policycosts_writeup.Rmd
exploring graphical representations for policycosts

``` {r setupsource}
rm(list=ls())   
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
```

```{r}
opts_knit$set(upload.fun = socialR::notebook.url)
opts_chunk$set(dev.args=list(bg="transparent"))
theme_set(theme_bw())
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
```

```{r}
delta <- 0.05     # economic discounting rate
OptTime <- 50     # stopping time
gridsize <- 50   # gridsize (discretized population)
sigma_g <- 0.2    # Noise in population growth
reward <- 0       # bonus for satisfying the boundary condition
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() 1 # measurement noise, none, but part of simulation routine
z_i <- function() 1
f <- BevHolt                # Select the state equation
pars <- c(1.5, 0.05)             # parameters for the state equation
K <- (pars[1] - 1)/pars[2]  # Carrying capacity (for reference 
xT <- 0                     # boundary conditions
x0 <- K
profit <- profit_harvest(price = 10, c0 = 30, c1 = 0)
x_grid <- seq(0.01, 1.2 * K, length = gridsize)  
h_grid <- seq(0.01, 0.8 * K, length = gridsize)  
````

```{r}
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g )
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime, xT, 
                     profit, delta, reward=reward)
````

``` {r fees}
L1 <- function(c2) function(h, h_prev)  c2 * abs(h - h_prev) 
free_increase <- function(c2) function(h, h_prev)  c2 * abs(min(h - h_prev, 0)) # increasing harvest is free
free_decrease <- function(c2) function(h, h_prev)  c2 * max(h - h_prev, 0) # decreasing harvest is free
fixed <-  function(c2) function(h, h_prev) c2 * as.numeric( !(h == h_prev) )
L2 <- function(c2) function(h, h_prev)  c2 * (h - h_prev) ^ 2
````

Solve the policy cost for the specified penalty function

``` {r L2}
L2_policy <- optim_policy(SDP_Mat, x_grid, h_grid, OptTime, xT, 
                    profit, delta, reward, penalty = L2(0.7586))
````

```{r fixed}
fixed_policy <- optim_policy(SDP_Mat, x_grid, h_grid, OptTime, xT, 
    profit, delta, reward, penalty = fixed(9.103))
```

``` {r L1}
L1_policy <- optim_policy(SDP_Mat, x_grid, h_grid, OptTime, xT, profit, delta, reward, penalty = L1(1.52))
````

```{r free_increase}
free_increase_policy <- optim_policy(SDP_Mat, x_grid, h_grid, OptTime, xT, profit, delta, reward, penalty =  free_increase(3.04))
````

```{r free_decrease}
free_decrease_policy <- optim_policy(SDP_Mat, x_grid, h_grid, OptTime, xT, profit, delta, reward, penalty = free_decrease(3.04))
````

``` {r simulate_policy}
sims <- list(
  L1 = simulate_optim(f, pars, x_grid, h_grid, x0, 
                      L1_policy$D, z_g, z_m, z_i, 
                      opt$D, profit=profit, penalty=L1(1.52), seed=111), 
  L2 = simulate_optim(f, pars, x_grid, h_grid, x0, 
                      L2_policy$D, z_g, z_m, z_i, 
                      opt$D, profit=profit, penalty=L2(0.7586), seed=111),
  fixed = simulate_optim(f, pars, x_grid, h_grid, x0, 
                         fixed_policy$D, z_g, z_m, z_i, 
                         opt$D, profit=profit, penalty=fixed(9.103), seed=111),
  increase = simulate_optim(f, pars, x_grid, h_grid, x0, 
                            free_decrease_policy$D, z_g, z_m, z_i, 
                            opt$D, profit=profit, penalty= free_increase(1.52), seed=111),
  decrease = simulate_optim(f, pars, x_grid, h_grid, x0, 
                            free_decrease_policy$D, z_g, z_m, z_i, 
                            opt$D, profit=profit, penalty= free_decrease(1.52), seed=111)
)
````

Make data tidy (melt), fast (data.tables), and nicely labeled.

``` {r tidy}
dat <- melt(sims, id=names(sims[[1]]))  
dt <- data.table(dat)
setnames(dt, "L1", "penalty_fn") # names are nice
````

# Plots

``` {r}
p0 <- ggplot(dt) +
  geom_line(aes(time, alternate), col="grey20", lwd=1) +
  geom_line(aes(time, fishstock, col=penalty_fn, lty=penalty_fn))+ 
  labs(x="time", y="stock size", title = "Stock Dynamics")
p0
```

```{r}
p1 <- ggplot(dt) +
  geom_line(aes(time, alternate), col="grey20", lwd=1) +
  geom_line(aes(time, fishstock), col=rgb(0,0,1,.8)) + facet_wrap(~penalty_fn) + 
  labs(x="time", y="stock size", title = "Stock Dynamics")
p1
```

```{r}
p2 <- ggplot(dt) +
  geom_line(aes(time, harvest_alt), col="grey20", lwd=1)  +
  geom_line(aes(time, harvest), col=rgb(0,0,1,.8)) + 
  facet_wrap(~penalty_fn) + 
  labs(x="time", y="havest intensity (fish taken)", title = "Harvest Policy Dynamics")
p2
````