### 2012-12-03 inst/examples/policycosts/policycosts_writeup_c0_30.Rmd
run on zero

```{r cache-options, include=FALSE}
opts_chunk$set(cache=TRUE, cache.path="policycosts-cache-c0-30/")
```

```{r}
profit <- profit_harvest(price = 10, c0 = 30, c1 = 0)
c2 <- seq(0, 20, length.out = 40)
```

```{r run, results="hide"}
knit("policycosts.Rmd", "policycosts_c0_30.md")
```

## Results

```{r }
ggplot(dat, aes(c2, (npv0-value)/npv0, col=variable)) + geom_point() + geom_line()
````

```{r}
as.table(apples)
```

``` {r }
p0
```

```{r fig.width=12}
p1 
p2
```