### 2012-12-04 inst/examples/policycosts/policycosts_writeup_c0_30.Rmd
example c0_30 from farm (images)

```{r cache-options, include=FALSE}
opts_chunk$set(cache=TRUE, cache.path="policycosts-cache-c0-30/", fig.path="figure/c30-")
```

```{r}
price = 1
c0 = 30
c1 = 0
```

```{r run, results="hide"}
knit("policycosts.Rmd", "policycosts_c0_30.md")
```

## Results

```{r }
apples_plot
````

```{r}
apples
```

```{r fig.width=12}
p1 
p2
```