### 2012-12-04 inst/examples/policycosts/policycosts_writeup_c0_30.Rmd
use price=10 for better scaling

```{r cache-options, include=FALSE}
opts_chunk$set(cache=TRUE, cache.path="policycosts-cache-c0-30/", fig.path="figure/c30-")
```

```{r price}
price = 10
c0 = 30
c1 = 0
```

```{r run, results="hide", dependson="price"}
knit("policycosts.Rmd", "policycosts_c0_30.md")
```

## Results

```{r dependson="run" }
apples_plot
````

```{r dependson="run"}
apples
```

```{r fig.width=12, dependson="run"}
p1 
p2
```