### 2013-11-05 inst/examples/policycosts/plot_table.Rmd
plottable

```{r}
opts_knit$set(upload.fun = socialR::flickr.url)
```

```{r}
require(ggplot2)
df <- read.csv("output.csv")
df <- df[2:8]

require(reshape2)
who <- c("penalty_fn", "ignore_fraction", "assume_fraction", "sigma_g", "reduction")
df2 <- melt(df, id=who)
who <- c("penalty_fn", "sigma_g", "reduction", "variable", "value")
df3 <- melt(df2, id=who)
names(df3) <- c("penalty_fn", "sigma_g", "reduction", "variable", "value", "fraction_variable", "fraction")

df3$reduction <- as.factor(df3$reduction)
#ggplot(df3) + geom_line(aes(sigma_g, value, col=penalty_fn, lty=variable)) + facet_wrap(~reduction) 
ggplot(df3) + geom_line(aes(sigma_g, value, col=reduction, lty=variable)) + facet_wrap(~penalty_fn) 
```

```{r results="asis"}
print(xtable::xtable(df3), type="html")
```