### 2013-11-08 inst/examples/policycosts/errors_table.Rmd
updates

``` {r libraries, message=FALSE, warning=FALSE}
rm(list=ls())
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
```

```{r cache-options, include=FALSE}
opts_chunk$set(cache=TRUE, cache.path="errors_table/", tidy=FALSE)
```

```{r}
start <- Sys.time()
```

``` {r defs}
compute_error_table <- function(r = 2, sigma = 0.2){
reduction_range = c(.1, .2, .3)
reduction <- reduction_range[r]

price = 10
c0 = 0
profit <- profit_harvest(price = price, c0 = c0, c1 = 0)
c2 <- seq(0, 50, length.out=100)

seed <- 123                 # Random seed (replicable results)
delta <- 0.05               # economic discounting rate
OptTime <- 20               # stopping time
gridsize <- 50              # grid size for fish stock and harvest rate (discretized population)
sigma_g <- sigma              # Noise in population growth
reward <- 0                 # bonus for satisfying the boundary condition
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() 1         # No measurement noise, 
z_i <- function() 1         # No implemenation noise
f <- BevHolt                # Select the state equation
pars <- c(1.5, 0.05)        # parameters for the state equation
K <- (pars[1] - 1)/pars[2]  # Carrying capacity (for reference 
xT <- 0                     # boundary conditions
x0 <- K
x_grid <- seq(0.01, 1.2 * K, length = gridsize)  
h_grid <- seq(0.01, 0.8 * K, length = gridsize)

sigma_g <- 0.3
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g )
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime, xT, 
                     profit, delta, reward=reward)

tmp <- lapply(1:100, function(i){ 
  set.seed(i)
  ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, z_m, z_i)
  })
tmp <- melt(tmp, id = names(tmp[[1]]))
sum(tmp$profit)

L1 <- function(c2) function(h, h_prev)  c2 * abs(h - h_prev) 
fixed <-  function(c2) function(h, h_prev) c2 * as.numeric( !(h == h_prev) )
L2 <- function(c2) function(h, h_prev)  c2 * (h - h_prev) ^ 2
none <- function(h, h_prev)  0

reduction_table <- read.csv('reduction_table.csv', row.names=1)
apples <-   reduction_table[[r]]
names(apples) <- rownames(reduction_table)

L2_policy <- optim_policy(SDP_Mat, x_grid, h_grid, OptTime, xT, 
                   profit, delta, reward, penalty = L2(apples[["L2"]]))
L1_policy <- optim_policy(SDP_Mat, x_grid, h_grid, OptTime, xT, 
                   profit, delta, reward, penalty = L1(apples[["L1"]]))
fixed_policy <- optim_policy(SDP_Mat, x_grid, h_grid, OptTime, xT, 
                   profit, delta, reward, penalty = fixed(apples[["fixed"]]))

reps <- 1:100
names(reps) = paste("rep", 1:100, sep="_")
seeds <- 1:100
sims <- list(
  L1 = lapply(reps, function(x) simulate_optim(f, pars, x_grid, h_grid, x0, 
                      L1_policy$D, z_g, z_m, z_i, 
                      opt$D, profit=profit, penalty=L1(apples[["L1"]]), seed=seeds[x])), 
  L2 = lapply(reps, function(x) simulate_optim(f, pars, x_grid, h_grid, x0, 
                      L2_policy$D, z_g, z_m, z_i, 
                      opt$D, profit=profit, penalty=L2(apples[["L2"]]), seed=seeds[x])),
  fixed = lapply(reps, function(x) simulate_optim(f, pars, x_grid, h_grid, x0, 
                         fixed_policy$D, z_g, z_m, z_i, 
                         opt$D, profit=profit, penalty=fixed(apples[["fixed"]]), seed=seeds[x]))

)

#Make data tidy (melt), fast (data.tables), and nicely labeled.
dat <- melt(sims, id=names(sims[[1]][[1]]))  
dt <- data.table(dat)
setnames(dt, "L2", "replicate") # names are nice
setnames(dt, "L1", "penalty_fn") # names are nice

# Profit when accounting for penalty when present
optimal_cost <- dt[, sum(profit_fishing - policy_cost), by=penalty_fn ] 
# Profit when ignoring penalty when present
ignore_when_present <- dt[, sum(profit_fishing_alt - policy_cost_alt), by=penalty_fn] 
# Profit when assuming penalty when it is absent
assume_when_absent <- dt[, sum(profit_fishing), by=penalty_fn]
# Profit when ignoring penalty when it is absent
optimal_free <- dt[, sum(profit_fishing_alt), by=penalty_fn]

# Normalize by the optimal 
ignore_fraction <- ignore_when_present$V1/optimal_free$V1 # common normalization
assume_fraction <- assume_when_absent$V1/optimal_free$V1
assume_when_absent <- cbind(assume_when_absent, assume_fraction = assume_fraction, normalize_optimal_free=optimal_free$V1, normalize_optimal_cost = optimal_cost$V1)
ignore_when_present <- cbind(ignore_when_present, ignore_fraction = ignore_fraction)

# Name and merge columns
setnames(ignore_when_present, "V1", "ignore_cost")
setnames(assume_when_absent, "V1", "assume_cost")
error_costs = merge(ignore_when_present, assume_when_absent, "penalty_fn")

optimal_cost$V1/optimal_free$V1 # actually realized

error_costs <- cbind(error_costs, sigma_g = sigma_g, reduction = reduction)

error_costs 
}

```

```{r compute_table}
sigmas <- c(0.01, 0.1, 0.2)
reductions <- 1:3
out <- lapply(sigmas, function(s){
  lapply(reductions, function(r){
    compute_error_table(r = r, sigma = s)    
  })
})

who <- names(out[[1]][[1]])
df <- melt(out, id=who)              
df
```

```{r results="asis"}
#df <- df[1:5]
library(xtable)
print(xtable(df), type="html")
writeLines(print(xtable(df)), "output.md")
system("pandoc output.md -o output.pdf")
write.csv(df, "output.csv")
```

```{r}
Sys.time() - start 
```