### 2014-11-30 manuscripts/manuscript.Rmd
revising text

---
title: Optimal management of a stochastically varying population when policy adjustment is costly
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTKEcon
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul Armsworth 
    affiliation: UTKEEB

address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research,
      110 Shaffer Rd,
      Santa Cruz, CA 95050, USA 
  - code: UCD
    address: Department of Environmental Science and Policy, University of California, Davis
  - code: melbourne
    address: School of Botany, University of Melbourne, Australia 
  - code: UTKEEB
    address: Department of Ecology and Evolutionary Biology, University of Tennessee, Knoxville
  - code: UTKEcon
    address: Department of Economics, University of Tennessee, Knoxville

abstract: |
          Ecological systems are dynamic and policies to manage them need to
          respond to that variation. However, policy adjustments will sometimes
          be costly, which means that fine-tuning a policy to track variability in
          the environment very tightly will only sometimes be worthwhile. We use a
          classic fisheries management question -- how to manage a stochastically
          varying population using annually varying quotas in order to maximize
          profit  -- to examine how costs of policy adjustment change optimal
          management recommendations. Costs of policy adjustment (here changes in
          fishing quotas through time) could take different forms. For example,
          these costs may respond to the size of the change being implemented or
          there could be a fixed cost any time a quota change is made. We show
          how different forms of policy costs have contrasting implications for
          optimal policies. Though it is frequently assumed that costs to adjusting
          policies will dampen variation in the policy, we show that certain cost
          structures can actually increase variation through time. We further show
          that failing to account for adjustment costs has a consistently worse
          economic impact than would assuming these costs are present when they are not.

layout: review,12pt
linenumbers: true

bibliography: components/references.bib
csl: components/ecology.csl
documentclass: components/elsarticle

output:
  pdf_document:
    template: components/elsarticle.latex
    fig_caption: true
    keep_tex: true
---

```{r configure, include=FALSE, cache=FALSE}
#source("components/install.R")

library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)
# PDF-based figures
opts_chunk$set(dev='pdf')
fig.cache <- TRUE

library("rmarkdown")
library("pdgControl")
library("reshape2")
library("plyr")
library("ggplot2")
library("data.table")
library("pander")
library("cboettigR")
library("ggthemes")
library("snowfall")
library("tidyr")
library("dplyr")
options(digits=2)

theme_set(theme_tufte(base_size=12) + theme(axis.line = element_line()))
```

```{r analysis, include=FALSE}
source("components/analysis.R")
```

Introduction
============

Ecosystems are dynamic and exhibit rich patterns of variability in
both time and space [@Durrett1994]. In designing management policies for ecosystems,
managers need to decide how much of that variation to respond to. Managers
could try to track variations in ecosystem dynamics very closely by
setting policies that are extremely responsive to the environment. Many
theoretical studies that seek to identify optimal policies for exploited
populations and communities adopt this approach and largely ignore the
challenges that would be involved in implementing such recommendations
[e.g., @Reed1979; @Neubert2003; @Sethi2005; @Halpern2011]. However, 
the policy process can often be much more sluggish
to respond to variations in ecosystem dynamics [@Walters1978; @Armsworth2010]. 
Moreover, stakeholders impacted by ecosystem management
may prefer some stability and not want to deal with continually changing
management recommendations [@Biais1995; @Armsworth2003;
@Patterson2007a; @Patterson2007b; @Sanchirico2008].
In other words, whatever gains are available from fine-tuning a
policy prescription to reflect environmental variation more closely
should be traded off against potential costs associated with the more
interventionist approach to management this would require.

```{r Figure_1, fig.cap="Estimated stock abundances (top panels) show much greater variation than the corresponding harvest policies (bottom).  Left panels: ICCAT Stock assessments and Total Allowable Catch (TAC, before adjustments for national underages and overages) for bluefin tuna, 1987 - 2007. Right panels: SEDAR stock assessments and annual quotas for south Atlantic snowy grouper.", fig.width=6, fig.height=4, dev=c("pdf", "svg", "png"), cache=FALSE}
tuna <- read.csv("components/data/tuna.csv")
grouper <- read.csv("components/data/grouper.csv")

ex <- rbind(cbind(tuna, species="tuna"), cbind(grouper, species="snowy grouper"))
ex_long <- gather(ex, variable, value, stock, quota)

ggplot(ex_long, aes(year,value)) + 
  geom_line() + geom_point() + 
  facet_wrap(variable~species, scales="free") + 
  ylab("value (mt)") + theme_bw()
```

To illustrate this tradeoff, we use an example from fisheries
management. Figure 1, top-left panel, shows a time series for the
estimated population size, represented as spawning stock biomass, of
west Atlantic bluefin tuna (_Thunnus thynnus_, henceforth bluefin) from
a recent stock assessment [@ICCAT20XX].  The bottom-left panel 
shows the catch quota (Total Allowable Catch or TAC, before being adjusted
for overages or underages in the catch of each country in the previous year)
for the stock as set by the relevant
management agency [@ICCAT20XX]. Despite the estimated population size
declining by XX% between YYYY and ZZZZ, the quota was not changed during
this period. Instead, the quota only changed occasionally and in between
times was left unaltered. Fishery management decisions regarding this
species can be highly contentious (@Safina1998, @Sissenwine1998,
@Porch2005). Moreover, the stock is fished by fleets from many nations
with quotas being set by a multilaterial management agency through a
process of negotiation. As such, we might reasonably anticipate that
for this species there could be substantial transaction costs involved
in reaching agreement over any quota change, which could contribute to
the observed quota stability. Such patterns of variable stocks managed
under far less variable quotas are extremely common; for instance, 
in the management of south Atlantic snowy grouper (_Epinephelus niveatus_) 
shown in the right-hand panels [@SEDAR2013]. More generally, reviews by @Biais1995 and
@Patterson2007a document many cases where changes in catch
quotas that a management agency set were more modest than changes that
would be recommended just by considering variations in stock abundance.

We use a classic fisheries management question to examine how accounting
for costs of policy adjustment can change optimal policies (see also
@Ludwig1980, @Feichtinger1994, @Wirl1999). We focus on how
harvest quotas for a stochastically varying fish population can be
chosen to maximize the net present value of a fishery. Our formulation
and solution method largely follow Reed's classic treatment on this
question, a treatment repeated widely in bioeconomic textbooks. We 
note that while later work has extended this treatment to deal with 
a variety of other issues (e.g. @Sethi2005; @Singh2006; @McGough2009),
we start from the classical model for simplicity of presentation and analysis.

With his formulation, @Reed1979 showed that a constant escapement policy could be
optimal under certain conditions. Such a policy involves choosing annual
quotas that are perfectly responsive to recruitment variation in a fish
stock. In poor recruitment years, the quota is set to zero and no fishing
is allowed. But any time there is a good recruitment pulse, a quota is
set that allows the fishery to exactly compensate through harvesting,
thereby maintaining the optimal escapement level. However, in that
analysis, Reed did not account for any costs of policy adjustment, which
for such a responsive management strategy potentially could be large.

As the examples in Figure 1 make clear, management policies will rarely be
as responsive as this constant escapement policy assumes. In this paper,
we explore conditions under which this lack of responsiveness may be a
rational response on the part of managers. Specifically, we consider a
case where managers seek to balance the benefits in terms of increased
profits from fishing from more finely tracking recruitment variations
with the growing costs associated with adjusting policies quickly to do
so. The policy adjustment costs involved could reflect pure administrative
transaction costs or preferences held by fishermen, fish processing plants
or other stakeholders, for less variable quotas. In seeking to account
for these policy adjustment costs, we recognize that we do not know just
what functional form they should take and that it will require a lot of
empirical work to estimate that. Therefore, we scope three candidate
functional forms that represent different assumptions about how these
costs operate to determine whether the results we obtain are sensitive
to such differences.

Methods
=======

Fish population dynamics (state equation)
-----------------------------------------

We will assume Beverton-Holt dynamics with multiplicative environmental noise

\begin{equation}
N_{t+1} = Z_t \frac{A (N_t - h_t)}{1 + B (N_t - h_t)}, \label{eq:state_equation}
\end{equation}

where $N_t$ the stock size, $h_t$ the harvested level, 
$Z_t$ gives the stochastic shocks, which we
assume are lognormally distributed and $A$ and $B$ are positive
constants.

We assume managers set an annual quota for harvesting $h_t$ (the control
variable) after observing the stock size that year $N_t$ (the state
variable), but while still being uncertain about future environmental
conditions and stock sizes. Through time, this gives a time path of
management actions ${\bf h}=(h_1,h_2,\dots)$ that depend on the stock
sizes that were observed (a state dependent control rule).

We assume that managers choose annual quotas to maximize the expected
net present value (NPV) of the fishery. We take as a base case the
situation where there are no costs associated with policy adjustment and
the managers' objective is

\begin{equation}
\max_{{\bf h}}\mathbf{ E} ( NPV_{0} )=\max_{{\bf h}} \sum_0^{T_{\mathrm{max}}} \mathbf{E} \left( \displaystyle \frac{\Pi_0(N_t,h_t)}{(1+\delta)^{t+1}} \right) \label{eq:objective}
\end{equation}

where $\mathbf{E}$ is the expectation operator, $\delta>0$ is the
discount rate and $\Pi_0$ is the net revenue from operating the fishery
in a given year. In this base case, we assume the annual dockside (that is,
before accounting for any adjustment costs that may be incurred) net revenue
from fishing is

\begin{equation}
\Pi_0(N_t,h_t) = p h_t -  c_0 E_t \label{eq:annual_netrevenue}
\end{equation}

where $E_t$ represents fishing effort. We assume catch is proportional
to stock size and effort expended fishing that year, $h_t = q E_t N_t$ and
constant $q>0$ is the catchability coefficient. In
Eqn. \eqref{eq:annual_netrevenue}, $p$ is the price per unit harvest and
$c_0$ the cost of fishing and, to simplify the presentation of results,
we assume that these are constants with $p>0$ and $c_0>0$.

To simplify presentation of the results, we will illustrate cases where the
growth parameters are chosen so that the intrinsic rate of increase of
the fish population is 1 and the equilibrium biomass for the equivalent
deterministic model without harvesting is 10; (specifically, $A=$ `r pars[1]` and
$B=$ `r pars[2]` in Eqn \eqref{eq:state_equation}. To characterize environmental variability, we
assume multiplicative shock $Z_t$ is distributed log-normally with 
log standard deviation $\sigma_g$ = `r sigma_g`. In addition, we
show cases where $p=$ `r price`, $\delta=$ `r delta` and $c_0=$ `r c0`.  
In the supplement we illustrate that the qualitative patterns observed
here are not sensitive to the specific choices for these parameter values.

Taken together this objective function \eqref{eq:objective} and the state
equation \eqref{eq:state_equation} define a stochastic dynamic programming
problem that we solve using backwards recursion via Bellman's equation.
We denote the resulting state-dependent, optimal control as $h_0^{\ast}$.
We solve this problem on a finite time horizon of $T =$ `r OptTime` using
value iteration [@Mangel1988; @Clark2000].

Note that this problem is much larger computationally than the classic 
formulation of this stochastic dynamic programming problem for a single stock.
Whereas the classic problem considers each of $Q$ possible discrete quotas
for $S$ possible stock values at each time $t$, for a search space of size $S \times Q$
per timestep;  this formulation must also consider all possible values of quota
in the _previous_ time step: a total of $S \times Q \times Q$ configurations.

R code for implementing the dynamic programming algorithm in this context is
provided as a supplementary R package to the paper. Scripts for replicating 
all of the analysis presented here are also provided in the Supplementary Material.

<!-- Cite & link to Zenodo DOI here(?)-->

Costs of policy adjustment 
--------------------------

We compare this base case to three alternative problem formulations,
each reflecting different plausible functional forms that costs of
policy adjustment could take. In each, we assume managers can adjust the
quota set in the fishery $h_t$ in a given year and that any policy
adjustment costs are associated with changes to this control variable.
In each case, we assume there is no cost in initially setting the harvest
policy at time 0.

First we assume that policy adjustment costs are directly proportional
to the magnitude of the change in policy being proposed, such that
larger changes to annual harvesting quotas incur greater policy
adjustment penalties. Specifically, we replace $\Pi_0$ in the $NPV_0$
equation with

\begin{equation}
\Pi_{1}(N_t,h_t, h_{t-1}) = \Pi_0 - c_1  |  h_t - h_{t-1} | \,.
\label{eq:Pi_1}
\end{equation}

Next we continue to assume that policy adjustment costs depend on the
magnitude of the change in policy being proposed. However, we consider a
case where this dependence is nonlinear with big changes in policy being
disproportionately expensive:

\begin{equation}
\Pi_{2}(N_t,h_t, h_{t-1}) = \Pi_0 - c_2 (  h_t - h_{t-1})^2 \,.
\label{eq:Pi_2}
\end{equation}

Finally, we assume instead that there is a fixed cost associated with
any adjustment of policy, regardless of how large that adjustment might
be

\begin{equation}
\Pi_{3}(N_t,h_t, h_{t-1}) = \Pi_0 - c_3 (1-\mathbf{I}(h_t, h_{t-1}))  \,,
\label{eq:Pi_3}
\end{equation}

where indicator function $\mathbf{I}(h_t, h_{t-1})$ takes the value 1 if
$h_t=h_{t-1}$ and zero otherwise. For each $\Pi_i$, we can then define a
new objective function $NPV_i$ similar to that in Eqn. \eqref{eq:objective}.

We note that the fixed cost has conceptual analogues to the set-up cost of
@Reed1974 and @Spulber1982 but here the fee is associated with a change
in harvesting ($h_t\neq h_{t-1}$) and not with harvesting *per se*
($h_t>0$) as in those studies.

Taken together with the state equation Eqn. \eqref{eq:state_equation}, each
of these new objective functions defines a different stochastic dynamic
programming problem. Again, we solve them numerically using backwards
recursion. To include costs of policy adjustment, we expand the state
space to include both the current stock size and the management action
taken on the previous time step $(N_t,h_{t-1})$. For each new objective
function $\max NPV_i$, we denote the corresponding optimal control
policy $h_i^*$.

In the Supplementary Material, we compare our results with 
more conventional fisheries economic formulation in which additional
costs are applied to the control variables themselves, as opposed to
adjustments to the controls: $\Pi_4 (N_t,h_t) = p h_t -  c_0 E_t-c_4E_t^2$. 
Additional costs of this form tend to have a smoothing effect on optimal quotas, but also change
the long-term average stock size or quota size that is optimal.

<!-- State conclusion of this supplement here! -->

Comparing apples to apples
--------------------------

Before we can compare the policy dynamics resulting from each of these
functional froms, we must be careful to ensure that we are only varying
the _structure_ of the adjustment penalty and not the overall magnitude.
At the extremes, a large enough adjustment cost can ensure a constant 
policy, regardless of the structure; while a small enough one can be ignored
regardless.  Defining this calibration correctly is thus key to the entire
analysis.  Fortunately, the optimal control framework provides a completely
natural way to make this comparison: in terms of the _economic value_
(measured by $NPV$) of the stock in each case.  We will calibrate this relative
to economic value of the stock when adjustments to the policy are free ($NPV_0$).

Each policy adjustment cost function is characterized in terms of
a cost coefficient $c_i$. However, $c_i$ takes different units for each
functional form. Therefore, if seeking to compare the relative effect of
each penalty function on optimal management, it is unclear what
parameter values should be used. To address this issue, we calibrate
choices of $c_i$ so that each has a comparable impact on the optimal NPV
fishery.

Figure 2 shows the calibration graphically. Each curve plots the change
in maximum expected $NPV$ for a given penalty function as the penalty cost
parameter is increased. In each case, the figure shows maximum expected
NPV with policy adjustment costs as a proportion of the maximum expected
NPV available in the basic problem $NPV_0({\bf h_0^*})$ without policy
adjustment costs. The $c_i$ values are given in Table 1.

```{r table_1, results="asis"}
pandoc.table(print_npv)
```

```{r Figure_2, fig.cap="Expected net present value by functional form of penalty. Horizontal axis shows the coefficient $c_i$ governing the magnitude of the policy cost, while vertical axis shows fraction of the maximum expected net present value dissipated by the policy cost, $\\tfrac{NPV_0 - NPV_i}{NPV_0}$.  The horizontal line indicates a value of the stock that is reduced by 25% from the maximum expected value in the absence of policy adjustment costs, $NPV_0({\\bf h_0^*})$. Selecting the coefficient $c_i$ corresponding to this value in each functional form allows us to make consistent comparisons across the different functional forms of policy costs.", dependson="quadcosts", fig.width=5, fig.height=3, cache = fig.cache, dev=c("pdf", "svg", "png")}

relabel <- c(L1 = substitute(paste(Pi[1])),
             L2 = substitute(paste(Pi[2])),
             fixed = substitute(paste(Pi[3])))

ggplot(fees, aes(c2, (npv0-value)/npv0, lty=variable, color=variable)) + 
  geom_line() + 
  geom_hline(aes(yintercept=reduction), linetype=4) + 
  xlab("Penalty coefficient") +
  ylab("Reduction in net present value") + 
  scale_linetype_discrete(labels=relabel, breaks = c('L1', 'L2', 'fixed')) + 
  scale_color_discrete(labels=relabel, breaks = c('L1', 'L2', 'fixed'))

write.csv(dt, "components/data/100reps.csv")
write.csv(fees, "components/data/figure2.csv")

```

To compare the impact of penalty functions on optimal management across
the different functional forms, we select penalty cost coefficients that
induce the same reduction in maximum expected NPV. For example, the
dashed vertical lines in Figure 2 map the needed cost coefficient $c_i$,
for each penalty function, such that the fishery is worth `r 100*(1-reduction)`% of its
unconstrained value when optimally managed. We focus our examination of the models on how
the type and severity of policy adjustment costs affects the optimal management strategy.

Results
=======

Effect of policy adjustment costs on optimal quotas and stock sizes 
-------------------------------------------------------------------

```{r Figure_3, dependson="tidy", fig.cap="Example realization of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=5, fig.width=8, cache = fig.cache, dev=c("pdf", "svg", "png")}

labeller <- function(variable,value){
    return(relabel[paste(value)])
}

dt3 <- dt %>% 
  filter(replicate=="rep_17" & time < 20) %>%
  select(time, harvest, harvest_alt, penalty_fn) %>%
  gather(variable, value, -time, -penalty_fn)

ggplot(dt3, aes(time, value, col=variable, lty=variable)) +
  geom_line(lwd=1) +
  facet_grid(penalty_fn~., labeller = labeller) + 
  labs(x="time", y="stock size", title = "Example Harvest Dynamics")  +
  scale_color_discrete(labels=c("With Penalty", "Reed Solution"), name="") +
  scale_linetype_discrete(labels=c("With Penalty", "Reed Solution"), name="")

write.csv(dt3, "components/data/figure3.csv")
```

Figure 3 illustrates how the different forms of adjustment cost can impact
the optimal harvest policy. Corresponding stock sizes can be seen in the
supplement.  Each panel is generated against the same sequence of 
environmental variability so that they can be compared directly.  The harvest policy chosen
shows a systematic deviations from the penalty free optimum, depending on
the nature of the cost.  In each case, the optimal solution without any
adjustment cost is shown with the policy induced by optimization
under the given adjustment penalty (equivalent to a `r 100*reduction`%
reduction in maximum expected NPV) overlaid in blue.

The first panel shows a typical pattern resulting from linear adjustment costs ($\Pi_1$).
The optimal policy tends to avoid very small policy adjustments, resulting in
periods of a constant policy followed by sudden bursts of adjustment.
This results in a relatively step-like policy pattern. In contrast, the
second formulation (quadratic costs, $\Pi_2$) disproportionately penalizes large policy
adjustments. The corresponding optimal policy is typified by the middle panel, responding
to each of the fluctuations in stock, made by the cost-free policy, but with smaller 
magnitude response than the equivalent cost-free optimal solution. This results
in a smoother $h_t$ curve, one that undershoots the larger oscillations
seen in the cost-free optimum in favor of a policy that changes
incrementally each year. Finally, the optimal policy for the third,
'fixed fee' formulation ($\Pi_3$) only makes large adjustments, as one
might expect, because the magnitude of the adjustment made is not
reflected in the resulting penalty.

These patterns are consistent across stochastic replicates over a range
of penalty magnitudes. The comparisons shown in Figure 3 are for one 
realization (i.e. a particular sequence of random number draws representing
environmental variability) and are made for one particular magnitude of 
policy adjustment costs (Calibrated to equal to `r reduction *100`% of the 
maximum expected net present valuein the base case ($NPV_0({\bf h_0^*})$,
as discussed in Figure 2, permitting comparisons between different forms).
To demonstrate this, we solved for the optimal policy under each of the
three adjustment cost scenarios for `r length(c0)` different penalty 
coefficients. For each resulting policy, we then simulated 50 stochastic
replicates of the stock dynamics managed under that policy.  To summarize
the patterns shown in Figure 3 and characterize the impact of increasingly
large fraction of the economic value (NPV) being consumed by policy
adjustment costs, we examine the response of several summary statistics
to increasing adjustment penalties (as a fraction of the adjustment-free value)
in Figure 4.  We show the impact of policy adjustment costs on the
variance & autocorrelation of the harvests through time.

```{r Figure_4, fig.cap="Variance and autocorrelation of harvest dynamics, as a function of the penalty coefficient.  Adjustment penalty is calculated as percentage of the maximum expected $NPV$ in the basic case with no policy adjustment costs.", fig.height=8, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png")}

library("dplyr")

f4df <- filter(figure4_df, timeseries=='harvest')
ggplot(f4df, aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn)) + 
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  coord_cartesian(xlim = c(0, .3)) + 
  facet_wrap(~ statistic, scale="free_y") + 
  scale_color_discrete(labels=relabel, name="Penalty function") + 
  scale_fill_discrete(labels=relabel, name="Penalty function") +
  xlab("Penalty size (as fraction of NPV0)")

write.csv(f4df, "components/data/figure4.csv")
```

Despite summarizing across such a large ensemble, Figure 4 confirms the
patterns we observed in the individual realizations shown in Figure 3.
For example, when small policy adjustments cost little but large
adjustments are expensive ($\Pi_2$), we see the smoothing signal that we
might have expected (see also @Ludwig1980 for this particular case). As
cost penalties increase in severity (moving right along the horizontal axis,
larger fractions of the economic value are consumed by adjustment costs), the
variance in quotas through time decreases and the autocorrelation of
quotas through time increases. 
When adjustment costs are zero, all policies are slightly negatively autocorrelated:
this reflects the fact that a high harvest year is usually followed by a low harvest year
that allows the stocks to recover. However, as adjustment costs increase, the quadratic
policy tends to undershoot when stocks jump to very high or low levels, adjusting to compensate
over multiple years, as seen in Figure 3.  As a result, autocorrelation increases.

Interestingly, as suggested by the realization in Fig. 3, including a
fixed cost ($\Pi_1$) of policy adjustment increases the variation in harvests
through time. This is the opposite of a smoothing effect.

Finally, the case where policy adjustment costs scale linearly with the
size of the adjustment ($\Pi_1$) appear to be something of a
middle of the road strategy, in that increasing the severity of policy
adjustment costs. To reveal the
particular impact of policy adjustment costs of this type requires a
more targeted summary statistic. Specifically, for each run we
calculated the frequency with which the optimal policy involved
maintaining a positive quota across multiple time steps unaltered. This
type of policy is arguably the most commonly observed behavior in TAC
management, but is one that is very rarely observed to be part of the
optimal management strategy in the basic model without policy adjustment
costs (Eqn. 3) or when optimizing against $\Pi_2$ or $\Pi_3$.

```{r}
fraction_no_shift <- function(x){

noshift <- x[2:length(x)] - x[1:length(x)-1] == 0
strictly_positive <- (x > 0.01 )[-1] # drop first point, 
who <- noshift & strictly_positive # True whenever consecutive pairs did not shift and the second in the pair was nonzero harvest
 sum(who)/length(who)
}

shifts <- dt[, fraction_no_shift(harvest), by=c("penalty_fn", "replicate")]
ave <- shifts[, mean(V1), by=penalty_fn]

no_shifts <- ave$V1
names(no_shifts) <- ave$penalty_fn
```

While still not common for the particular parameter combinations we examine,
We find that positive unaltered quotas through time are much more likely
to occur when optimizing against $\Pi_1$ where policy adjustment costs
scale linearly with the size of proposed policy changes: over the 100 replicates
time series the harvest policy is strictly positive and identical in consecutive intervals
only `r no_shifts[["L2"]] * 100`% of the time for quadratic costs $\Pi_2$, compared with
`r no_shifts[["L1"]] * 100`% with linear costs $\Pi_1$ and `r no_shifts[["fixed"]]*100`%
for fixed costs $\Pi_3$. Moreover, these occurrences increase in frequency as the severity of
these costs ($c_i$) increases.

Consequences of policy adjustment costs 
---------------------------------------

```{r histogram-calc}
profits <- dt[, sum(profit_fishing), by=c('penalty_fn', 'replicate') ]
costs <- dt[, sum(policy_cost), by=c('penalty_fn', 'replicate') ]
reed_profits <- dt[, sum(profit_fishing_alt), by=c('penalty_fn', 'replicate') ]
reed_costs <- dt[, sum(policy_cost_alt), by=c('penalty_fn', 'replicate') ]
setnames(profits, "V1", "profits")
setnames(reed_profits, "V1", "profits")

Reed <- cbind(reed_profits, costs = reed_costs$V1, Assumption = "Reed") 
Adj <- cbind(profits, costs = costs$V1, Assumption = "Adjustment penalty")

hist_dat <- melt(rbind(Adj, Reed), id=c("penalty_fn", "replicate", "Assumption"))
````

```{r}
fig5_caption <- paste("Distribution of profits from fishing and costs paid to adjustment, under linear costs ($\Pi_1$). (Alternate penalties can be found in the supplementary materials)  Distributions are shown when managing a system in which the adjustment costs present compared to the same costs and profits of managing the system using a Reed model, which ignores adjustment costs, and thus tends to pay higher fees for adjustment.  In contrast, the difference in profit after accounting for costs is only marginally less than the value derived by managing the population to optimize profits when ignoring adjustment costs (Reed model).")
```

```{r Figure_5, fig.cap=fig5_caption, fig.width=8, fig.height=6, dependson="histogram-calc", dev=c("pdf", "svg", "png"), cache=FALSE}

fig5_dat <- hist_dat %>% 
  filter(penalty_fn=="L1") %>%
  select(-penalty_fn) %>%
  unite(dist, Assumption, variable, sep = " : ")

ggplot(fig5_dat) + 
  geom_density(aes(value, fill=dist, color=dist), alpha=0.8)

write.csv(hist_dat, "components/data/figure5.csv")
```

Next we examine the consequences either of ignoring policy adjustment
costs when they are present or assuming they are present when they are
not. To do so, we run two kinds of simulations:

1. Simulations in which the optimal policy accounts for adjustment costs, 
   and those costs are present (in the form assumed),
2. Simulations in which the optimal policy _does not_ account for adjustment costs,

While the optimization guarentees that on average, the policy in the first simulation
should outperform that of the second, it does not tell us by how much.  These simulations
thus give us a lense to examine the magnitude of this effect.

We provide 100 replicates of each kind for each of the three forms
$\Pi_{1-3}$ (in which each ensemble of 100 replicates uses the same sequence of random seeds).
For each, we record (a) the dockside profits derived from fishing as well
as (b) any fees paid to make adjustments to the policy.  In the first kind of 
simulation, dockyard profits are almost less than in the second for a given stochastic
replicate, since the latter optimizes the expected dockside NPV.  However, the costs
paid for adjustment are almost always less in the first scenario as well, since
this solution explicitly tries to avoid higher adjustment costs.  We visualize this
pattern in the distributions shown in Figure 5 (focusing on $\Pi_1$ for simplicity,
as the pattern in $\Pi_2$ and $\Pi_3$ is similar, see Supplementary Material.)

Figure 5 shows that the difference in dockside revenue (labelled "profits" in the figure) 
between these two scenarios 
(green vs purple distributions) is very small, while a larger difference exists between
the costs paid for making policy adjustments (red vs blue). This is predicted by the optimization: 
the first scenario must outperform (or at least equal) the NPV of the second scenario
in aggregate, and as it cannot do so through higher profits it must do so through lower
costs.  Interestingly, for this set of parameters very little is lost in dockside revenue.
This suggests that the impact of assuming costs are present when they are not is rather small:
since the dockside revenue is none-the-less near optimal.  This also suggests that the 
converse error: ignoring costs when they are present, is more severe.

```{r Figure_6, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue), shown by penalty function over increasing magnitudes of cost.  All values are relative to the cost-free adjustment optimum, $NPV_0$.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. The severity of the latter mistake increases sharply as adjustment costs make a larger fraction of the total costs.", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}

who <- c("penalty_fn", "ignore_fraction", "assume_fraction", "reduction")
table1 <- arrange(error_df[who], reduction) 
names(table1) = c("penalty.fn", "ignoring", "assuming", "reduction")
table1_long <- melt(table1, id = c('penalty.fn', 'reduction'))
table1_long$reduction <- as.factor(table1_long$reduction)
table1_long <- subset(table1_long, reduction != "0.3")

ggplot(filter(table1_long, reduction == 0.25),
  aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, ncol=2) + 
  scale_x_discrete(labels = c('L1' = substitute(paste(Pi[1])),
                              'L2' = substitute(paste(Pi[2])),
                              'fixed' = substitute(paste(Pi[3])))) + 
  xlab("Penalty function")
write.csv(table1_long, "components/data/figure6.csv")
```

We visualize the differences between these two errors directly in Figure 6. Subtracting revenue from
adjustment costs and considering only the mean across the replicates, we can dervive the expected $NPV$ 
in each case.  To standardize against a common baseline, we will express this value as a fraction of 
the cost-free optimum, $NPV_0$. Note that we have calibrated the optimal solutions to anticipate that
`r reduction*100`% of their $NPV$ would go into adjustment costs. When these adjustment costs are not 
actually present (blue bars, Fig 6), we thus expect a $NPV$ higher than the anticipated `r (1 -reduction)*100`%
but not quite as high as the optimal policy for that scenario (100%).  When adjustment costs are present
but the policy assumes that they are not (red bars), we would expect the $NPV$ to be less than `r (1 - reduction) * 100`%
of $NPV_0$ (since that's as good as the optimal solution for this scenario can do).

Figure 6 shows that across each penalty form $\Pi{1-3}$, the economic impact
of ignoring adjustment costs when they are present is higher than accounting
for them when they are absent. This holds even after accounting for effect 
described above: the red bars are consistently well below (`r (1-reduction)`),
while the blue bars are barely below 1.
We also see that the impacts on net present value of the fishery of
incorrectly assuming or ignoring policy adjustment costs are most severe
when these costs are assumed to scale quadratically with the size of the
policy change being implemented ($\Pi_2$).
In the Supplementary Materials we show that pattern holds across different
severity of adjustment costs (robust to choice of $c_i$), though the larger
the adjustment cost, the greater the difference between these two errors.

Note that this analysis does not consider the possibility that an adjustment cost
is present, but of a different functional form than the policy assumes.  Not surprisingly,
the consequences of managing under the wrong functional form can be as or more severe than
simply ignoring adjustment costs in the optimization all-together.  We illustrate several
of these mismatch errors in the Supplementary Material.

Discussion
==========

  Policy-makers managing ecological systems that vary in space and time
must evaluate how much of that variation to reflect in management
recommendations. Fine-tuning a policy to respond to frequent variations
in ecological dynamics may incur increased transaction costs associated
with constantly revisiting past policy decisions. A more pragmatic
approach would be one that balances benefits from responding to frequent
variations in ecological conditions with the increased transaction costs
involved. As a first step towards exploring these ideas, we revisited
a classic problem from bioeconomics concerning the optimal management
of a fish stock subject to stochastically varying recruitment [@Reed1979; @Clark2010]. 
We examined how optimal policy recommendations, here
annual catch quotas, changed when accounting for costs associated with
policy adjustment and what the implications of following these policy
recommendations would be for the exploited population. We also compared
how the value of the fishery is affected by managers either under-
or over-estimating the importance of policy adjustment costs of this type.

Estimating policy adjustment costs and how they respond to the size of
proposed policy changes would be empirically challenging. Recognizing
this fact, in our analyses, we compared different plausible forms that
these costs might take. We compared two cost structures where we assumed
the magnitude of policy adjustment costs increased with the magnitude
of the change in policy being proposed with an alternative formulation
in which we assumed there was a fixed cost associated with making any
change to current policy. While each formulation clearly provides only a
very phenomenological representation of the policy setting environment,
we believe that each enables us to explore meaningful differences in how
policy adjustment costs might operate. That being said, in the real-world
we might expect the different types of adjustment cost to operate in
combination.

The biggest differences between the representations of policy adjustment
cost that we consider are between those that have a smoothing effect
on annual quotas and those that do not. The few past studies that
incorporate policy adjustment costs in models of fisheries management
[@Ludwig1980; @Feichtinger1994; @Wirl1999] assumed costs of
policy adjustment increased as a quadratic function of the magnitude of
the policy change being proposed (our $\Pi_2$ formulation) In
effect, this means that small changes to annual quotas incur little
extra cost, but large changes to annual quotas become disproportionately
expensive to make. Including costs of this form smoothes inter-annual
variation in the catch quotas that would be recommended. Also, the catches and remaining stock sizes
corresponding to optimal management become more autocorrelated in time,
because it takes the fishery longer to harvest down peaks in abundance
that follow large recruitment pulses. The effects of smoothing here are
similar to those predicted when assuming the cost per unit effort is
increasing in the amount of effort expended (e.g., @Lewis1981, @McGough2009; 
see Supporting Information for a summary of the relevant
results) as opposed to associating extra costs with changes to policy
*per se*. Smoothing effects of this type are what one might have
expected when including policy adjustment costs.

Standing in sharp contrast to these smoothing predictions is our
finding that including policy adjustment costs can actually _increase_ the
variability of quotas through time if there is a fixed cost associated
with making any changes to current policy (e.g. costs of running relevant
stakeholder meetings and public consultations on proposed policy changes,
$\Pi_3$). With this formulation, as stock sizes vary in response to
recruitment, the fishery manager must balance the cost of brokering
a policy change with the cost in forgone fisheries revenue from not
responding to favorable recruitment pulses. The optimal policy involves
ignoring small variations in recruitment, but then assigns a larger quota
when particularly strong recruitment years arise than would have been
the case in the absence of policy adjustment costs. The other functional
form we consider, in which the costs of policy adjustment scale linearly
with the size of the adjustment ($\Pi_1$), has less obvious effects on
optimal policies. But interestingly, it is the formulation that
most frequently produces stretches of strictly positive but unchanging
quotas of the type most commonly encountered in real-world applications.

That the different representations of policy adjustment costs result in
such different dynamics suggests that researchers constructing fisheries
economics models should proceed cautiously when choosing how to represent
these costs. However, in our own experience, we have found that while
modeling studies sometimes mention policy adjustment costs when motivating
model assumptions (references), they rarely provide much justification
for the choice of functional form used or test the sensitivity of any
conclusions drawn to alternative specifications. Indeed, we were surprised
to find such clear differences between the functional forms, because
we anticipated that the limiting case of increasing policy adjustment
costs within each functional form should be the same, namely a constant
annual quota that does not change through time (open-loop control).

We also compared the efficiency costs that would result from failing
to account for policy adjustment cost if they are present with those
involved in assuming them when in fact they are absent. The results
of this comparison were not sensitive to the particular form of policy
adjustment costs assumed. Instead, we always found the efficiency costs
of ignoring policy adjustment costs when they were present to be much
larger than the efficiency costs of assuming policy adjustment costs
applied when they were in fact absent. Policy adjustment costs affect
the overall value of the fishery in two ways here. First, there is the
direct cost associated with each quota change. Second there is a cost in
foregone revenue from missed catches when not following what would be the
optimal policy if quotas were free to track recruitment variability. Our
finding that it is more costly to ignore policy adjustment costs when
they are present arises because the first, more direct, cost contribution
here is larger than the second.

Assuming an adjustment cost exists when in fact it is absent is not the
same as using the wrong functional form when it is present.  Here, our
results showed that choice of the functional form makes a significant
difference.  We show it can be better to ignore policy costs than to derive a 
policy based on assuming the wrong functional form.  This result further
underscores the importance of modelers using caution when seeking to account
for these costs. Assuming an arbitrary form in order to capture the influence
of adjustment costs is thus unlikely to be instructive.

As with any modeling analysis, our formulation makes many assumptions. For
example, we assume that the fishery in question is being optimally managed
by a policy-maker who acts as the 'sole owner' of the stock. This approach
is different to models that assume fisheries are not well-managed,
e.g. by assuming open access conditions,
or ones that derive policy recommendations endogenously by modeling
interactions between different stakeholders [@Kaitala1993; @Laukkanen2003].
As such, we anticipate that our approach will be more relevant to
some fisheries, particularly domestic fisheries in developed countries
that are subject to relatively strong regulatory regimes, than to others
(e.g., artisanal fisheries that are subject to weaker regulation). To
focus on the effects of policy adjustment costs, we focused narrowly
on the basic model specification of @Reed1979 and examined how the
predictions of this classic problem were changed by introducing policy
adjustment costs of different forms. However, there have been many
elaborations on Reed's basic approach that increase the realism of
the optimization models involved by relaxing other assumptions (see for
example @Sethi2005; @Singh2006; @McGough2009).

One obvious research avenue suggested by our models is that of empirically
estimating costs of policy adjustment. A direct estimation approach
could quantify some sources of policy adjustment costs, e.g. costs to
processing plants that arise from having more variable catches. However,
other more intangible sources of policy adjustment costs, e.g. preferences
of policy-makers or different stakeholders for less variable quotas,
might be missed. An alternative, more holistic, approach would be to
apply revealed preference methods to fisheries management agencies
themselves, in the tradition of [@McFadden1975; @McFadden1976]. Such an analysis
would involve comparing quotas that were set relative to stock sizes as
they were estimated at the time each management decision was taken to
try to infer what objective managers were maximizing.

A worthwhile modeling extension suggested by our results would be to
examine the implications of policy adjustment costs for risks of stock
collapse. Examining risks of stock collapse is not possible with the
Reed model-formulation that we followed here but would be worthwhile
in light of our finding that the variance and autocorrelation in stock
sizes through time can be affected when accounting for costs of policy
adjustment. Both are properties known in other modeling settings to
be associated with changes to the risk of extinction or of transitions
between alternative stable states [@Scheffer2009; @Boettiger2013].

<!-- Comment on increasing variance or increasing autocorrelation from changes in policy costs / optimal management-->

When facing highly variable ecological systems, how often should
natural resource managers respond? A highly interventionist strategy
would track ecosystem variation very closely. Alternatively, a manager
might choose only to take action or change policy only when conditions
look very different to those previously experienced. We took a modeling
approach to begin to explore these ideas. Specifically, we focused on
a well-known problem from fisheries management and examined how optimal
management recommendations changed when we accounted for costs associated
with frequently changing management decisions. While we focused on a
fisheries context, our findings would be relevant to many other settings
where natural resource managers revisit management decisions through
time in light of ecological variability, including game management,
management of in-stream flow rates, fire management, habitat restoration,
and managing for endangered species. In some ways, the analyses that
we present can also be thought of as providing a temporal counter-part
to discussions about the optimal spatial scale over which ecosystem
management should be conducted.

Acknowledgements
================

This project arose out of a NIMBioS working group, "Pretty Darn Good Control".
The authors acknowledge helpful discussions and input from working group
co-organizers Megan Donahue and Carl Toews, and participants Marie-Josee Fortin, 
Dan Ryan, Frank Doyle, Claire Paris, Iadine Chades and Mandy Karnauskas.   
We also acknowledge the following support: NSF Grant DBI-1306697 (CB), ...

# References

# Supplementary material #

##  Impact on the stock dynamics ##

```{r StockDynamics, dependson="tidy", fig.cap="Example realization of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=5, fig.width=8, cache = fig.cache, dev=c("pdf", "svg", "png")}

labeller <- function(variable,value){
    return(relabel[paste(value)])
}

dt_s <- dt %>% 
  filter(replicate=="rep_17" & time < 20) %>%
  select(time, fishstock, alternate, penalty_fn) %>%
  gather(variable, value, -time, -penalty_fn)

ggplot(dt_s, aes(time, value, col=variable, lty=variable)) +
  geom_line(lwd=1) +
  facet_grid(penalty_fn~., labeller = labeller) + 
  labs(x="time", y="stock size", title = "Example Stock Dynamics")  +
  scale_color_discrete(labels=c("With Penalty", "Reed Solution"), name="") +
  scale_linetype_discrete(labels=c("With Penalty", "Reed Solution"), name="")

```

```{r StockStats, fig.cap="Variance and autocorrelation of stock dynamics, as a function of the penalty coefficient.  Adjustment penalty is calculated as percentage of the maximum expected $NPV$ in the basic case with no policy adjustment costs.", fig.height=8, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png")}

stock_stats <- filter(figure4_df, timeseries=='stock')
ggplot(stock_stats, aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn)) + 
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  coord_cartesian(xlim = c(0, .3)) + 
  facet_wrap(~ statistic, scale="free_y") + 
  scale_color_discrete(labels=relabel, name="Penalty function") + 
  scale_fill_discrete(labels=relabel, name="Penalty function") +
  xlab("Penalty size (as fraction of NPV0)")

```

Recall that in our model definition, stock sizes are measured before harvest ($h_t$ being determined after observing stock size
$x_t$) and thus show less influence of the policy cost than the harvest itself.

## Additional stochastic realizations ##

Figure 3 of the main text shows a single stochastic realization.  Here we provide a sample of further realizations (including both stock and harvest dynamics).  Full

```{r Figure_S2, dependson="tidy", fig.cap="Multiple realizations of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=10, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png"), cache=FALSE}

labels <- c(harvest = substitute(paste("Harvest Policy under ", paste(Pi[1-3]))), 
            harvest_alt = substitute(paste("Harvest under ", paste(Pi[0]), " (Reed)")),
            fishstock = substitute(paste("Stock dynamics under under ", paste(Pi[1-3]))), 
            alternate = substitute(paste("Stock under ", paste(Pi[0]), " (Reed)")))
labeller <- function(variable,value){
if (variable=='penalty_fn') {
    return(relabel[paste(value)])
  } else {
    return(value)
  }
}

ex <- dt %>% 
  filter(replicate %in% paste0("rep_", 16:20) & time < 20) %>%
  select(time, fishstock, harvest, alternate, harvest_alt, replicate, penalty_fn) %>%
  gather(variable, stock, -time, -replicate, -penalty_fn)

ggplot(ex) + 
  geom_line(aes(time, stock, col=variable), lwd=1) + 
  scale_color_discrete(labels=labels, name="") +  
  facet_grid(replicate ~ penalty_fn, labeller=labeller)

write.csv(ex, "components/data/figure_s2.csv")
```

## Economic penalties using quadratic costs ##

```{r quadraticcosts}

source("components/quad_costs.R")
```

```{r Figure_S1, dev=c('pdf', 'svg'), cache=FALSE}

labels <- c(harvest = substitute(paste("Harvest Policy under ", paste(Pi[4]))), 
                                harvest_alt = substitute(paste("Harvest under ", paste(Pi[0]), " (Reed)")))

ex <- dt_quad %>% 
  filter(replicate=="rep_17" & time < 20) %>%
  select(time, harvest, harvest_alt) %>%
  gather(variable, stock, -time) 
ggplot(ex) + 
  geom_line(aes(time, stock, col=variable), lwd=1) + 
  scale_color_discrete(labels=labels, 
                       name="")

```

```{r quadshiftrate}
shifts <- dt[, fraction_no_shift(harvest), by=c("replicate")]
ave <- shifts[, mean(V1)]
```

Like the $\Pi_2$ adjustment cost based penalty, this quadratic cost based penalty shows periods of contant, strictly positive policy only `r ave * 100`% of the time, and exhibits the same kind of smoothing of more small adjustments that is seen in $\Pi_2$ scenario.

<!-- Figure 5 extended across -->

```{r Figure_S3, fig.cap="Distribution of profits from fishing and costs paid to adjustment.  Distributions are shown when managing a system in which the adjustment costs present are determined by the model indicated, and compared to the same costs and profits of managing the same system using a Reed model, which ignores adjustment costs, and thus tends to pay much higher fees for adjustment.  In contrast, the difference in profit after accounting for costs is only marginally less than the value derived by managing the population to optimize profits when ignoring adjustment costs (Reed model).", fig.width=8, fig.height=6, dependson="histogram-calc", dev=c("pdf", "svg", "png"), cache=FALSE}

assume <- levels(as.factor(hist_dat$Assumption))
labeller <- function(variable,value){
if (variable=='penalty_fn') {
    return(relabel[paste(value)])
  } else {
    return(paste(value))
    #return(assume[paste(value)])
  }
}

ggplot(hist_dat) + 
  geom_density(aes(value, fill=variable, color=variable), alpha=0.8)+
  facet_grid(Assumption~penalty_fn, labeller = labeller)

write.csv(hist_dat, "components/data/figure5.csv")
```

```{r Figure_S4, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue), shown by penalty function over increasing magnitudes of cost.  All values are relative to the cost-free adjustment optimum, NPV0.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. The severity of the latter mistake increases sharply as adjustment costs make a larger fraction of the total costs.", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}

ggplot(table1_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, ncol=2) + 
  scale_x_discrete(labels = c('L1' = substitute(paste(Pi[1])),
                              'L2' = substitute(paste(Pi[2])),
                              'fixed' = substitute(paste(Pi[3])))) + 
  xlab("Penalty function")
write.csv(table1_long, "components/data/figure6.csv")
```

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure S5 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10% (panel A)
less than optimal free based on the $c_i$ calibration. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

For fixed costs ($\Pi_3$), the reverse is true.  Because fixed costs increase the
volatility of the optimal solution while $\Pi_2$ costs decrease it, simply
ignoring the cost of adjustment all together results in a higher value
than either assuming costs are quadratic ($\Pi_2$) when they are fixed ($\Pi_3$),
or assuming they are fixed when they are quadratic.  Scenarios in which true costs are quadratic (regardless of what
cost structure is assumed by the policy) do significantly worse than
those that are not (i.e. all cases that assume quadratic but actually
have linear or fixed costs).

These patterns hold across the different size reductions shown (from 10%
to 25%, panels a-d respectively), though net present value becomes negative under the quadratic
form for the larger penalties as costs paid for adjustments exceed
profits derived from harvest.  (Though optimal solutions will always
avoid negative net present values, these scenarios in which policies are
applied under conditions other than assumed in the optimality calculation
have no such guarantee.) For each scenario, results are averaged over
100 replicates.

```{r Figure_S5, fig.cap="Mismatches. X axis labels indicate the model used to derive the policy, followed by the model used to simulate the penalty. Blue bars show the net present value of that scenario.  Compare to red bars, which show the consequence of using a policy that ignores costs all together in the same scenario (i.e. costs as indicated by second code), rather than assuming the policy indicated by the first code in the pair.  All values are scaled to a fraction of the adjustment free net present value.", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

ggplot(filter(table2_long, reduction==0.25), 
  aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
write.csv(table2_long, "components/data/figure7.csv")
```

<!--

```{r Figure_S5, fig.cap="Effect of varying severity of adjustment costs on mismatches. Successive panels show increasing reduction of the net present value by increasing penalty coefficients: e.g. a policy in the first panel would lose only 10% (panel A) of its net present value to adjustment costs if it had the correct form of those costs, as determined by the choice of penalty coefficient.  Plots structured as in figure 7, which corresponds to the fourth panel", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}
ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
write.csv(table2_long, "components/data/figure7.csv")
```

-->