### 2015-05-12 manuscripts/supplement.Rmd
additional runs at different NPV reduction calibration points

---
title: "Supplement to: Optimal management of a stochastically varying population when policy adjustment is costly"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTKEcon
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul R. Armsworth 
    affiliation: UTKEEB

address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research,
      110 Shaffer Rd,
      Santa Cruz, CA 95050, USA 
  - code: UCD
    address: Department of Environmental Science and Policy, University of California, Davis
  - code: melbourne
    address: School of Botany, University of Melbourne, Australia 
  - code: UTKEEB
    address: Department of Ecology and Evolutionary Biology, University of Tennessee, Knoxville
  - code: UTKEcon
    address: Department of Economics, University of Tennessee, Knoxville

layout: review,12pt
linenumbers: true

bibliography: components/references.bib
csl: components/ecology.csl
documentclass: components/elsarticle

output:
  pdf_document:
    template: components/elsarticle.latex
    fig_caption: true
---

```{r configure, include=FALSE, cache=FALSE}
library("methods")
library("knitr")
basename <- "supplement"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)
# PDF-based figures
opts_chunk$set(dev='pdf')
fig.cache <- TRUE

library("rmarkdown")
library("pdgControl")
library("reshape2")
library("plyr")
library("ggplot2")
library("data.table")
library("pander")
library("cboettigR")
library("ggthemes")
library("snowfall")
library("tidyr")
library("dplyr")
options(digits=2)

theme_set(theme_tufte(base_size=12) + theme(axis.line = element_line()))
```

\renewcommand{\thefigure}{S\arabic{figure}}
\appendix

# Supplementary material #

## Code and data for methods and analysis ##

All code and data used in this analysis can be found under the project's GitHub repository: [https://github.com/cboettig/pdg_control](https://github.com/cboettig/pdg_control). A snapshot of this repository representing the precise versions of the code used in this analysis has been permanently archived in the scientific software and data repository Zenodo, managed by CERN; and should be referenced or cited using its assigned digital object identifer, doi:xxx.  (This DOI is also clearly linked from the GitHub repository).

### Organization of code

The project repository is organized as an R package.  The intent is not to provide a general-purpose software tool for this kind of analysis, but rather to rely on a standardized directory structure for representing complex projects. It is our hope that this facilates understanding, reproducibility, and further exploration of these results sensu [Gentleman & Temple Lang 2007](http://biostats.bepress.com/bioconductor/paper2/), but caution that additional software engineering would be required before any of this could be applied in a different context.  The functions provided are largely specific to the analyses presented here and may undocumented or incompletely documented.

The source documents for both the manuscript and supplement are also found in the code repository using the dynamic document `.Rmd` format which embeds all of the code required for the analysis; see [http://rmarkdown.rstudio.com/](http://rmarkdown.rstudio.com/).

Within the `.Rmd` files themselves are found only the blocks of code required to generate each of the figures shown in the manuscript.  The analysis is sourced in from an external R script, `analysis.R` by a code block at the beginning of the `.Rmd` file.  This script in turn relies on more modular functions defined in the `R/` directory that define the stochastic dynamic programming optimization algorithms, simulation routines, and so forth.

### Data

All empirical or simulated data shown in the figures can also be found as tabular .csv files in this repository.

### Dependencies and machine images

##  Impact on the stock dynamics ##

Figure S1 shows the corresponding stock dynamics for the replicates illustrated in Figure 3 of the main text.  The impact of the adjustment cost on the stock dynamics is less dramatic than it is on the harvest dynamics, due to the stochastic growth in stock sizes and the sequence in which the steps are applied: stock values are assessed without error, then an optimal harvest is determined (using the pre-computed optimal policy based on both the stock assessment and the previous harvest, to account for possible adjustment costs), and then the stock that escapes harvesting recruits stochastically to set the population size for the next year.

Though this results in an effect that is less pronounced, the stock dynamics of Figure S1 reflect the same general trends observed in Main text Figure 3; e.g. the fixed fee $\Pi_3$ results in the most exaggerated fluctuations in the stock sizes.  It is worth noting how changes in policy costs alone, such as introducing or rising fixed costs, can have important knock-on effects to population dynamics such as increased variance, a recent phenomenon observed in important fisheries that has received much attention elsewhere [@Anderson2008].

```{r Figure_S1, dependson="tidy", fig.cap="Example realization of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=5, fig.width=8, cache = fig.cache, dev=c("pdf", "svg", "png")}

labeller <- function(variable,value){
    return(relabel[paste(value)])
}

dt_s <- dt %>% 
  filter(replicate=="rep_17" & time < OptTime) %>%
  select(time, fishstock, alternate, penalty_fn) %>%
  gather(variable, value, -time, -penalty_fn) 
dt_s$penalty_fn <- factor(dt3$penalty_fn, levels=c("L1", "L2", "fixed"))

ggplot(dt_s, aes(time, value, col=variable, lty=variable)) +
  geom_line(lwd=1) +
  facet_grid(penalty_fn~., labeller = labeller) + 
  labs(x="time", y="stock size", title = "Example Stock Dynamics")  +
  scale_color_discrete(labels=c("With Penalty", "Reed Solution"), name="") +
  scale_linetype_discrete(labels=c("With Penalty", "Reed Solution"), name="")

write.csv(dt_s, "components/data/figureS1.csv")
```

```{r Figure_S2, fig.cap="Variance and autocorrelation of stock dynamics, as a function of the penalty coefficient.  Adjustment penalty is calculated as percentage of the maximum expected $NPV$ in the basic case with no policy adjustment costs.", fig.height=6, fig.width=4, cache=fig.cache, dev=c("pdf", "svg", "png")}

stats_df %>% 
  filter(variable != "cross.correlation") %>%
  separate(variable, c("measurement", "statistic"), sep="\\.") %>%
  filter(measurement == 'stock') %>% 
  spread(statistic, value) ->
  s2df

s2df$penalty_fn <- factor(f4df$penalty_fn, levels = c("L1", "L2", "fixed"))

p4 <- ggplot(s2df, aes(x = penalty_fraction, fill=penalty_fn, col=penalty_fn)) + 
  coord_cartesian(xlim = c(0, .3)) + 
  scale_color_discrete(labels=relabel, name="cost model") + 
  scale_fill_discrete(labels=relabel, name="cost model")

p4a <- p4 + 
  stat_summary(aes(y=autocorrelation),
               fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(aes(y=autocorrelation), fun.y = mean, geom = "line") + 
  xlab("")

p4b <- p4 + 
  stat_summary(aes(y=variance),
               fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(aes(y=variance), fun.y = mean, geom = "line") +
  xlab(expression(paste("Penalty size (as fraction of ", NPV[0], ")")))

legend <- gtable_filter(ggplot_gtable(ggplot_build(p4a)), "guide-box") 
grid.arrange(do.call(arrangeGrob, 
            lapply(list(p4a, p4b),
                   function(x) x + theme_bw() + theme(legend.position="none"))),
    legend,
    ncol = 2,
    widths = c(5,1))

write.csv(s2df, "components/data/figureS2.csv")

```

Recall that in our model definition, stock sizes are measured before
harvest ($h_t$ being determined after observing stock size $N_t$) and
thus show less influence of the policy cost than the harvest itself.

Figure S2 shows the same summary statistics illustrating the trends
of increasing adjustment costs on the _stock_ dynamics, instead of the
_harvest_ dynamics shown in Figure 4 of the main text.  Note that 
the variance in stock sizes under the quadratic penalty ($\Pi_2$) increases,
while in Figure 4 of the main text the corresponding variance in the
harvest decreases. Under costly adjustment, the undershooting behavior
of the $\Pi_2$ harvest policy cannot absorb as much of the natural 
fluctuation, resulting in increased variation in the stock sizes.

## Additional stochastic realizations ##

Figure 3 of the main text shows only a single stochastic realization.  Here in
Figure S3 we provide a sample of further realizations (including both
stock and harvest dynamics).

```{r Figure_S3, dependson="tidy", fig.cap="Multiple realizations of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=10, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png"), cache=FALSE}

labels <- c(harvest = substitute(paste("Harvest Policy under ", paste(Pi[1-3]))), 
            harvest_alt = substitute(paste("Harvest under ", paste(Pi[0]), " (Reed)")),
            fishstock = substitute(paste("Stock dynamics under under ", paste(Pi[1-3]))), 
            alternate = substitute(paste("Stock under ", paste(Pi[0]), " (Reed)")))
labeller <- function(variable,value){
if (variable=='penalty_fn') {
    return(relabel[paste(value)])
  } else {
    return(value)
  }
}

dt$penalty_fn <- factor(dt$penalty_fn, levels = c("L1", "L2", "fixed"))

ex <- dt %>% 
  filter(replicate %in% paste0("rep_", 16:20) & time < OptTime) %>%
  select(time, fishstock, harvest, alternate, harvest_alt, replicate, penalty_fn) %>%
  gather(variable, stock, -time, -replicate, -penalty_fn)

ggplot(ex) + 
  geom_line(aes(time, stock, col=variable), lwd=1) + 
  scale_color_discrete(labels=labels, name="") +  
  facet_grid(replicate ~ penalty_fn, labeller=labeller)

write.csv(ex, "components/data/figureS3.csv")
```

## Economic penalties using quadratic costs ##

Figure S4 shows that adding quadratic terms to the cost of fishing effort directly, as in @Lewis1981 or @McGough2009, results in a similar smoothing effect as we observe in quadratic adjustment costs ($\Pi_2$), which contrasts sharply with the patterns we observe in other functional forms.  This case corresponds to a profit function defined as:

$$\Pi_4(N_t,h_t) = p h_t -  c_0 E_t - c_4 E_t^2$$

Where $c_4$ is calibrated as in the case of the other coefficients to induce the fixed reduction of net present value relative to $NPV_0$.

```{r quadraticcosts, include=FALSE, message=FALSE}

source("components/quad_costs.R")
```

```{r Figure_S4, dev=c('pdf', 'svg'), cache=FALSE, fig.cap="Quadratic costs to fishing effort show a similar pattern of smoothing to that which we observe in the quadratic adjustment cost model."}

labels <- c(harvest = substitute(paste("Harvest Policy under ", paste(Pi[4]))), 
                                harvest_alt = substitute(paste("Harvest under ", paste(Pi[0]), " (Reed)")))
ex <- dt_quad %>% 
  filter(replicate=="rep_17" & time < OptTime) %>%
  select(time, harvest, harvest_alt) %>%
  gather(variable, stock, -time) 
ggplot(ex) + 
  geom_line(aes(time, stock, col=variable), lwd=1) + 
  scale_color_discrete(labels=labels, 
                       name="")

```

```{r quadshiftrate}
shifts <- dt[, fraction_no_shift(harvest), by=c("replicate")]
ave <- shifts[, mean(V1)]
```

Like the $\Pi_2$ adjustment cost based penalty, this quadratic cost based penalty shows periods of constant, strictly positive policy only `r ave * 100`% of the time, and exhibits the same kind of smoothing of more small adjustments that is seen in $\Pi_2$ scenario.

Ignoring adjustment costs is worst than wrongly assuming adjustment costs
--------------------------------------------------------------------------

Figure S5 extends Figure 5 of the main text across all three functional forms of adjustment costs, $\Pi_{1-3}$. To reduce over-plotting of Figure 5 we have also separated the distributions of dockside revenue and adjustment fees by scenario, such that the top row shows the distributions when the policy accounts for adjustment costs, and the bottom shows the policy when it does not (that is, the original @Reed1979 model applied to the scenario of costly adjustment).

```{r Figure_S5, fig.cap="Distribution of profits from fishing and costs paid to adjustment.  Distributions are shown when managing a system in which the adjustment costs present are determined by the model indicated, and compared to the same costs and profits of managing the same system using a Reed model, which ignores adjustment costs, and thus tends to pay much higher fees for adjustment.  In contrast, the difference in profit after accounting for costs is only marginally less than the value derived by managing the population to optimize profits when ignoring adjustment costs (Reed model).", fig.width=8, fig.height=6, dependson="histogram-calc", dev=c("pdf", "svg", "png"), cache=FALSE}

assume <- levels(as.factor(hist_dat$Assumption))
labeller <- function(variable,value){
if (variable=='penalty_fn') {
    return(relabel[paste(value)])
  } else {
    return(paste(value))
    #return(assume[paste(value)])
  }
}

ggplot(hist_dat) + 
  geom_density(aes(value, fill=variable, color=variable), alpha=0.8)+
  facet_grid(Assumption~penalty_fn, labeller = labeller)

write.csv(hist_dat, "components/data/figureS5.csv")
```

Figure S6 extends the results shown in Figure 6 across different severities of adjustment cost, totaling from 10% up to 25% of the $NPV$ of the stock. While the same qualitative pattern is observed in all cases, it is strongest in the higher penalties.

```{r Figure_S6, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue), shown by penalty function over increasing magnitudes of cost.  All values are relative to the cost-free adjustment optimum, NPV0.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. The severity of the latter mistake increases sharply as adjustment costs make a larger fraction of the total costs.", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}

table1_long$penalty.fn <- factor(table1_long$penalty.fn, levels = c("L1", "L2", "fixed"))

ggplot(table1_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, ncol=2) + 
  scale_x_discrete(labels = c('L1' = substitute(paste(Pi[1])),
                              'L2' = substitute(paste(Pi[2])),
                              'fixed' = substitute(paste(Pi[3])))) + 
  xlab("Penalty function")
write.csv(table1_long, "components/data/figureS6.csv")
```

Mismatched penalty functions
----------------------------

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure S7 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10% (panel A)
less than optimal free based on the $c_i$ calibration. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

For fixed costs ($\Pi_3$), the reverse is true.  Because fixed costs increase the
volatility of the optimal solution while $\Pi_2$ costs decrease it, simply
ignoring the cost of adjustment all together results in a higher value
than either assuming costs are quadratic ($\Pi_2$) when they are fixed ($\Pi_3$),
or assuming they are fixed when they are quadratic.  Scenarios in which true costs are quadratic (regardless of what
cost structure is assumed by the policy) do significantly worse than
those that are not (i.e. all cases that assume quadratic but actually
have linear or fixed costs).

```{r Figure_S7, fig.cap="Mismatches. X axis labels indicate the model used to derive the policy, followed by the model used to simulate the penalty. Blue bars show the net present value of that scenario.  Compare to red bars, which show the consequence of using a policy that ignores costs all together in the same scenario (i.e. costs as indicated by second code), rather than assuming the policy indicated by the first code in the pair.  All values are scaled to a fraction of the adjustment free net present value.", dev=c("pdf", "svg", "png"), fig.width=7}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

ggplot(filter(table2_long, reduction==reduction), 
  aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
write.csv(table2_long, "components/data/figureS7.csv")
```

These patterns hold across the different size reductions shown (from 10%
to 25%, Figure S8 panels a-d respectively), though net present value becomes negative under the quadratic
form for the larger penalties as costs paid for adjustments exceed
profits derived from harvest.  (Though optimal solutions will always
avoid negative net present values, these scenarios in which policies are
applied under conditions other than assumed in the optimality calculation
have no such guarantee.) For each scenario, results are averaged over
100 replicates.

```{r Figure_S8, fig.cap="Effect of varying severity of adjustment costs on mismatches. Successive panels show increasing reduction of the net present value by increasing penalty coefficients: e.g. a policy in the first panel would lose only 10% (panel A) of its net present value to adjustment costs if it had the correct form of those costs, as determined by the choice of penalty coefficient.  Plots structured as in figure 7, which corresponds to the fourth panel", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}
ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
```

## Robustness to reduction level

```{r}
source("components/vary_reduction.R")
```

Figures S9-S12 illustrate that the 
these qualitative patterns still hold at when the models are calibrated 
to a reduction in NPV of 10% or 30%, though naturally the magnitude
of the pattern is greater when the penalty is larger (e.g. 30% vs 10%).

```{r Figure_S9, dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE, fig.cap="Identical analysis to Figure 3 of the main text, but using a calibration corresponding to a 10% reduction in NPV.  This results in a less pronounced version of the same qualitative pattern."}
p3 <- ggplot(low$dt3, aes(time, value, col=variable, lty=variable)) +
  geom_line(lwd=1) +
  facet_grid(penalty_fn~., labeller = labeller) + 
  labs(x="time", y="stock size", title = "Example Harvest Dynamics")  +
  scale_color_discrete(labels=c("Adjustment penalty", "No penalty"), name="") +
  scale_linetype_discrete(labels=c("Adjustment penalty", "No penalty"), name="") + 
  theme_bw() + labs(title="example trajectories at 10% reduction in NPV")

p3
```

```{r Figure_S10, dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE, fig.cap="Identical analysis to Figure 3 of the main text, but using a calibration corresponding to a 30% reduction in NPV.  This results in a more pronounced version of the same qualitative pattern."}
p3 <- ggplot(high$dt3, aes(time, value, col=variable, lty=variable)) +
  geom_line(lwd=1) +
  facet_grid(penalty_fn~., labeller = labeller) + 
  labs(x="time", y="stock size", title = "Example Harvest Dynamics")  +
  scale_color_discrete(labels=c("Adjustment penalty", "No penalty"), name="") +
  scale_linetype_discrete(labels=c("Adjustment penalty", "No penalty"), name="") + 
  theme_bw() + labs(title="example trajectories at 30% reduction in NPV")
p3
```

```{r Figure_S11, dev=c("pdf", "svg", "png"), fig.width=5, fig.height=6, cache=FALSE, fig.cap="Identical analysis as in Figure 4 of the main text, but using a calibration to 10% reduction in NPV rather than 25%.  The qualitatively similar pattern helps illustrate that these results are not specific to the value of the reduction chosen in the main text."}
p4 <- ggplot(low$f4df, aes(x = penalty_fraction, fill=penalty_fn, col=penalty_fn)) + 
  coord_cartesian(xlim = c(0, .3)) + 
  scale_color_discrete(labels=relabel, name="cost model") + 
  scale_fill_discrete(labels=relabel, name="cost model")

p4a <- p4 + 
  stat_summary(aes(y=autocorrelation),
               fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(aes(y=autocorrelation), fun.y = mean, geom = "line") + 
  xlab("") + 
  labs(title="Summary statistics under a 10% reduction in NPV")

p4b <- p4 + 
  stat_summary(aes(y=variance),
               fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(aes(y=variance), fun.y = mean, geom = "line") +
  xlab(expression(paste("Penalty size (as fraction of ", NPV[0], ")")))

legend <- gtable_filter(ggplot_gtable(ggplot_build(p4a)), "guide-box")

grid.arrange(do.call(arrangeGrob, 
            lapply(list(p4a, p4b),
                   function(x) x + theme_bw() + theme(legend.position="none"))),
    legend,
    ncol = 2,
    widths = c(5,2))
```

```{r Figure_S12, dev=c("pdf", "svg", "png"), fig.width=5, fig.height=6, cache=FALSE, fig.cap="Identical analysis as in Figure 4 of the main text, but using a calibration to 30% reduction in NPV rather than 25%.  The qualitatively similar pattern helps illustrate that these results are not specific to the value of the reduction chosen in the main text."}
p4 <- ggplot(high$f4df, aes(x = penalty_fraction, fill=penalty_fn, col=penalty_fn)) + 
  coord_cartesian(xlim = c(0, .3)) + 
  scale_color_discrete(labels=relabel, name="cost model") + 
  scale_fill_discrete(labels=relabel, name="cost model")

p4a <- p4 + 
  stat_summary(aes(y=autocorrelation),
               fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(aes(y=autocorrelation), fun.y = mean, geom = "line") + 
  xlab("") + 
  labs(title="Summary statistics under a 30% reduction in NPV")

p4b <- p4 + 
  stat_summary(aes(y=variance),
               fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(aes(y=variance), fun.y = mean, geom = "line") +
  xlab(expression(paste("Penalty size (as fraction of ", NPV[0], ")")))

legend <- gtable_filter(ggplot_gtable(ggplot_build(p4a)), "guide-box")

grid.arrange(do.call(arrangeGrob, 
            lapply(list(p4a, p4b),
                   function(x) x + theme_bw() + theme(legend.position="none"))),
    legend,
    ncol = 2,
    widths = c(5,2))
```